Formal Analysis of EDHOC Key Establishment for Constrained IoT Devices

### 2022-07-18 formalization.tex
Changes to section 3

\subsection{Formal Model}\label{sec:threat-model}
\vnote{Can we add extra references here? SAPIC+ etc?} As in~\cite{Norr21}, we verify \mEdhoc{} in the Dolev-Yao model~\cite{DY83}.
%
This is a well-established model for symbolic verification of security protocols.
%
Messages are modelled as terms in an algebra, and the various cryptographic operations are assumed to be perfect, i.e. encrypted messages can only be decrypted with the correct key, no hash collisions exist etc.
%

@@

We now briefly describe \mTamarin{}, and how protocols are modelled in it.

%-------------------------------------------------------------------------- sub
\subsection{\mTamarin{}}
\label{sec:tamarin}
As in~\cite{Norr21}, we model and verify \mEdhoc{} in the symbolic model using \mTamarin{}~\cite{DBLP:conf/cav/MeierSCB13}, which is an interactive tool for the formal verification of security protocols.
%
Protocols are modelled in \mTamarin{} as multi-set rewrite rules which encode a transition relation.
%
The elements of these multi-sets, called facts, contribute to the global system state.

@@

An annotated multi-set rewrite rule is represented as $l \ifarrow[e] r$, where $l$ and $r$ are multi-sets, and $e$ is a multi-set of actions.
%
A sequence of actions yields a protocol execution trace.
%
Properties are defined as formulas in a first-order logic, and these formulas can be verified over execution traces.

@@

For syntactic sugar, \mTamarin{} allows the use of let-bindings and tuples. 
%
The rules modelling a Dolev-Yao adversary and the evolution of their knowledge as the protocol executes are built into \mTamarin.
%
However, we can extend and/or modify this behaviour by adding other rules as well.

@@

%%%%%%%%%%% START HERE

\subsection{Formalization of Properties}
\label{sec:desired-properties}
We use the \mTamarin{} verification
tool~\cite{DBLP:conf/cav/MeierSCB13} to encode the model and verify properties.
%
This tool uses a fragment of temporal first order logic to reason about
events and knowledge of the parties and of the adversary.
%
For conciseness we use a slightly different syntax than
that used by \mTamarin{}, but which has a direct mapping to \mTamarin{}'s logic.
%

Event types are predicates over global states of system execution.
%
Let $E$ be an event type and let $t$ be a timestamp associated with a point in a
trace.
%
Then $E^{t}(p_i)_{i\in\mathbb{N}}$ denotes an event of type $E$ associated with
a sequence of parameters $(p_i)_{i\in\mathbb{N}}$ at time $t$ in that trace.
%
In general, more than one event may have the same timestamp and hence
timestamps form a quasi order, which we denote by $t_1 \lessdot t_2$ when $t_1$
is before $t_2$ in a trace.
%
We define $\doteq$ analogously.
%
However, two events of the same type cannot have the same timestamp, so
$t_1 \doteq t_2$ implies $E^{t_1} = E^{t_2}$.
%
Two events having the same timestamp does not imply the there is a fork in the
trace, only that the two events happen simultaneously and that comparing their
timestamps results in that they are equal.
%
This notation corresponds to \mTamarin{}'s use of action facts
$E(p_i)_{i\in\mathbb{N}}@t$.
%

The event $\mK^t(p)$ denotes that the adversary knows a parameter $p$ at
time $t$.
%
Parameters are terms in a term algebra of protocol specific operations and
generic operations, e.g., tuples $\langle\cdot\rangle$.
%
Intuitively, $\mK^t(p)$ evaluates to true when $p$ is in
the closure of the
parameters the adversary observed from interacting with parties using the
protocol, under the Dolev-Yao message deduction operations and
the advanced adversary capabilities up until time $t$.
%
For a more precise definition of knowledge management, we refer to~\cite{DBLP:conf/cav/MeierSCB13}.
%
An example of a formula is
\[
    \forall t, k, k'\mLogicDot \mK^{t}(\langle k, k'\rangle)\ \rightarrow\ 
\mK^{t}(k) \land \mK^{t}(k'),
\]
expressing that if there is a time $t$ when the adversary knows the tuple
$\langle k, k'\rangle$, then the adversary knows both
$k$ and $k'$ at the same point in time.
%

An initiator $I$ considers the
protocol run started when it sends a message \mMsgone{} (event type \mIStart)
and the run completed after sending a message \mMsgthree{} (event type
\mIComplete).
%
Similarly, a responder $R$ considers the run started upon receiving
a \mMsgone{} (event type \mRStart), and completed upon receiving a \mMsgthree{}
(event type \mRComplete).
%

%----------------------------------------------------------------------- PFS
\subsubsection{Perfect Forward Secrecy (PFS)}
\label{sec:secrecy}
Informally, PFS captures the idea that session key material remains secret
even if a long-term key leaks in the future.
%
We define PFS for session key material \mSessKey{} as \mPredPfs{} in
Figure~\ref{fig:props}.
%

\begin{figure*}
\begin{align*}
    \mPredPfs \triangleq\ & \forall I, R, \mSessKey, t_2, t_3\mLogicDot
    \mK^{t_3}(\mSessKey)\  \land\ 
    (\mIComplete^{t_2}(I, R, \mSessKey)\, \lor\, \mRComplete^{t_2}(I, R, \mSessKey))
    \rightarrow\\
    &(\exists t_1\mLogicDot \mRevLTK^{t_1}(I) \land t_1 \lessdot t_2)
    \ \lor\ (\exists t_1\mLogicDot \mRevLTK^{t_1}(R) \land t_1 \lessdot t_2)
    \ \lor\ (\exists t_1\mLogicDot \mRevEph^{t_1}(R, \mSessKey))
    \ \lor\ (\exists t_1\mLogicDot \mRevEph^{t_1}(I, \mSessKey))
\end{align*}
\begin{align*}
    \mPredInjI \triangleq\ &
    \forall I, R, \mSessKey, S, t_2\mLogicDot \mIComplete^{t_2}(I, R, \mSessKey, S)
    \rightarrow\\
    &(\exists t_1\mLogicDot \mRStart^{t_1}(R, \mSessKey, S) \land t_1 \lessdot t_2)
    \land (\forall I' R' t_1' \mLogicDot \mIComplete^{t_1'}(I' , R', \mSessKey, S)
        \rightarrow t_1' \doteq t_1)
    \ \ \lor\ \ (\exists t_1\mLogicDot \mRevLTK^{t_1}(R) \land t_1 \lessdot t_2)
\end{align*}
\begin{align*}
    \mPredInjR \triangleq\ &
    \forall I, R, \mSessKey, S, t_2\mLogicDot \mRComplete^{t_2}(I, R, \mSessKey, S)
    \rightarrow\\
    &(\exists t_1\mLogicDot \mIStart^{t_1}(I, R, \mSessKey, S) \land t_1 \lessdot t_2)
    \land (\forall I' R' t_1' \mLogicDot \mRComplete^{t_1'}(I' , R', \mSessKey, S)
        \rightarrow t_1' \doteq t_1)
    \ \ \lor\ \ (\exists t_1\mLogicDot \mRevLTK^{t_1}(I) \land t_1 \lessdot t_2).
\end{align*}
\begin{align*}
    \mPredImpI \triangleq\ &
    \forall I, R, \mSessKey, S, t_1\mLogicDot \mIComplete^{t_1}(I, R, \mSessKey, S)
    \rightarrow\\
      &(\forall I', R', S', t_2\mLogicDot \mRComplete^{t_2}(I', R', \mSessKey, S') \rightarrow
             (I=I' \land R=R' \land S=S'))\\
      &\land (\forall I', R', S', t_1'\mLogicDot
        (\mIComplete^{t_1'}(I', R', \mSessKey, S') \rightarrow t_1' \doteq t_1
        )\\
      &\ \ \ \lor(\exists t_0\mLogicDot \mRevLTK^{t_0}(R) \land t_0 \lessdot t_1)
    \lor(\exists t_0\mLogicDot \mRevEph^{t_0}(R, \mSessKey))
    \lor(\exists t_0\mLogicDot \mRevEph^{t_0}(I, \mSessKey)).
\end{align*}
%
\caption{Formalization of security properties and adversary model.}
\label{fig:props}
\end{figure*}

The first parameter $I$ of the \mIComplete{} event represents the
initiator's identity,
and the second, $R$, represents that $I$ believes $R$ to be playing
the responder role.
%
The third parameter, \mSessKey{}, is the established session key material.
%
The parameters of the \mRComplete{} event are defined analogously.
%
Specifically, the first parameter of \mRComplete{} represents the identity of
whom $R$ believes is playing the initiator role.
%
The essence of the definition is that an adversary only knows \mSessKey{} if they
compromised one of the
parties long-term keys before that party completed the run, or if the adversary
compromised any of the ephemeral keys at any time after a party starts
its protocol run.
%
One way the definition slightly differs from the corresponding \mTamarin{} lemma
is that \mTamarin{} does not allow a disjunction on the left-hand side of an
implication in a universally quantified formula.
%
In the lemma, therefore, instead of the disjunction
$\mIComplete^{t_2}(I, R, \mSessKey)\, \lor\,  \mRComplete^{t_2}(I, R, \mSessKey)$,
we use a single action parametrized by $I$, $R$, and \mSessKey{} to signify that
\emph{either} party has completed their role.
%

%-------------------------------------------------------------------- InjAgree
\subsubsection{Authentication}
\label{sec:authenticationDef}
We prove two different flavors of authentication, the first being classical
\emph{injective agreement} following Lowe~\cite{DBLP:conf/csfw/Lowe97a}, and
the second being an implicit agreement property.
%
Informally, injective agreement guarantees to an initiator $I$ that whenever
$I$ completes a run ostensibly with a responder $R$,
then $R$ has been engaged in the protocol as a responder,
and this run of $I$ corresponds to a unique run of $R$.
%
In addition, the property guarantees to $I$ that the two parties agree on a set
$S$ of parameters associated with the run, including, in particular, the
session key material \mSessKey{}.
%
However, we will treat \mSessKey{} separately for clarity.
%
On completion, $I$ knows that $R$ has access to the session key material.
%
The corresponding property for $R$ is analogous.
%

Traditionally, the event types used to describe injective agreement are called
\emph{Running} and \emph{Commit}, but to harmonize the presentations of
authentication and PFS in this section, we refer to these event types as
\mIStart{} and \mIComplete{} respectively for the initiator, and
\mRStart{} and \mRComplete{} for the responder.
%
For the initiator role we define injective agreement by
\mPredInjI{} in Figure~\ref{fig:props}.

%
The property captures that for an initiator $I$, either the injective agreement
property as described above holds, or the long-term key of the believed
responder $R$ has been compromised before $I$ completed its role.
%
Had the adversary compromised $R$'s long-term key, they could have generated a
message of their liking (different to what $R$ agreed on), and signed this or
computed a $\mathit{MAC}_R$ based on \mPubE{I}, \mPriv{R} and their own chosen
ephemeral key pair $\langle\mPrivE{R},\ \mPubE{R}\rangle$.
%
This places no restrictions on the ephemeral
key reveals, or on the reveal of the initiator's long-term key.
%
For the responder we define the property \mPredInjR{} as in
Figure~\ref{fig:props}.
%

%------------------------------------------------------------- Implicit auth
Unlike PFS, not all \mEdhoc{} methods enjoy the injective agreement property.
%
Hence, we show for all methods a form of \emph{implicit agreement} on all the
parameters mentioned above.
%
We take inspiration from the computational model definitions of implicit
authentication, proposed by Guilhem~et~al.~\cite{DBLP:conf/csfw/GuilhemFW20}, to
modify classical injective agreement into an implicit property.
%
A small but important difference between our definition and theirs, is that
they focus on
authenticating a key and related identities, whereas we extend the more general
concept of agreeing on a set of parameters, starting from the idea of injective
agreement~\cite{DBLP:conf/csfw/Lowe97a}.
%
We use the term \emph{implicit} in this context to denote that a party $A$
assumes that any other party $B$ who knows the session key material \mSessKey{} must
be the intended party, and that $B$ (if honest) will also agree on a set
$S$ of parameters computed by the protocol, one of which is \mSessKey{}.
%
When implicit agreement holds for both roles, upon completion, $A$ is guaranteed
that $A$ has been or is engaged in exactly one protocol run with $B$ in the
opposite role, and that $B$ has been or will be able to agree on $S$.
%
The main difference to injective agreement is that $A$ concludes that if
$A$ sends the last message and this reaches $B$, then $A$ and $B$ have agreed
on $I$, $R$ and $S$.
%
While almost full explicit key authentication, as defined by
Guilhem~et~al.~\cite{DBLP:conf/csfw/GuilhemFW20}, is a similar property, our
definition does not require key confirmation, so our definition is closer to
their definition of implicit authentication.
%
In the \mTamarin{} model we split the property into one lemma for
$I$ (\mPredImpI{}) and one for $R$ (\mPredImpR{}) to save memory during
verification.
%
We show only the definition for $I$ in Figure~\ref{fig:props}, because it is
symmetric to the one for $R$.
%

For implicit agreement to hold for the initiator $I$, the ephemeral keys
can never be revealed.
%
Intuitively, the reason for this is that the implicit agreement relies on that
whomever knows the session key material is the intended responder.
%
An adversary with access to the ephemeral keys and the public keys of
both parties can compute the session key material produced by all methods.
%
However, the responder $R$'s long-term key can be revealed after $I$ completes
its run, because the adversary is still unable to compute $P_e$.
%
The initiator's long-term key can also be revealed at any time without affecting
$I$'s guarantee for the same reason.
%

%------------------------------------------------------- Agreed parameters
\subsubsection{Agreed Parameters}
\label{sec:agreedParams}
The initiator $I$ gets injective and implicit agreement guarantees on the
following partial set $S_P$ of parameters:
\begin{itemize}
    \item the roles played by itself and its peer,
    \item responder identity,
    \item session key material (which varies depending on \mEdhoc{} method),
    \item context identifiers \mCi{} and \mCr{}, and
    \item cipher suites \mSuites{}.
\end{itemize}
%
Because \mEdhoc{} aims to provide identity protection for $I$, there is no
injective agreement guarantee for $I$ that $R$ agrees on the initiator's
identity.
%
For the same reason, there is no such guarantee for $I$ with respect to
the $P_I$ part of the session key material when $I$ uses the \mStat{}
authentication method.
%
There is, however, an implicit agreement guarantee for $I$ that $R$ agrees on
$I$'s identity and the full session key material.
%
Since $R$ completes after $I$, $R$ can get injective agreement guarantees on
more parameters, namely also the initiator's identity and the full session key
material for all methods.
%
The full set of agreed parameters $S_F$ is $S_P \cup \{I, P_I\}$
when $P_I$ is
part of the session key material, and $S_P \cup \{I\}$ otherwise.
%

%------------------------------------------------------- Implied properties
\subsubsection{Inferred Properties}
From the above, other properties can be inferred to hold in our adversary model.
%
Protocols where a party does not get confirmation that their peer knows the
session key material may be susceptible to
\emph{Key-Compromise Impersonation (KCI)}
attacks~\cite{DBLP:conf/ima/Blake-WilsonJM97}.
%
Attacks in this class allow an adversary in possession of a party $A$'s secret
long-term key to coerce $A$ to complete a
protocol run believing it authenticated a certain peer $B$, but where $B$ did
not engage with $A$ at all in a run.
%
Because both our notions of agreement above ensure agreement on identities,
roles and session key material, all methods passing verification of those are
also resistant to KCI attacks.
%

If a party $A$ can be coerced into believing it completed a run with $B$, but
where the session key material is actually shared with $C$ instead, the 
protocol is vulnerable to an \emph{Unknown Key-Share (UKS)}
attack~\cite{DBLP:conf/ima/Blake-WilsonJM97}.
%
For the same reason as for KCI, any method for which our agreement
properties hold is also resistant to UKS attacks.
%

From the injective agreement properties it follows that each party is assured
the identity of its peer upon completion.
%
Therefore, the agreement properties also capture \emph{entity authentication}.
%

%---------------------------------------------------------------------------sub
\subsection{\mTamarin{} Encoding of \mEdhoc{}}
\label{sec:modeling}
%
We model all four methods of \mEdhoc{}, namely
\mSigSig, \mSigStat, \mStatSig{} and \mStatStat.
%
Because the methods share a lot of common structure, we derive
their \mTamarin-models from a single specification written with the aid of the
M4 macro language.
%
To keep the presentation brief, we only present the \mStatSig{} metohod, as it
illustrates the use of two different asymmetric authentication methods
simultaneously.
%
The full \mTamarin{} code for all models can be found at~\cite{edhocTamarinRepo}.
%
Variable names used in the code excerpts here are sometimes shortened compared
to the model itself to fit the paper format.
%

>>>>

\subsection{Formal Model}\label{sec:threat-model}
\vnote{Can we add extra references here? SAPIC+ etc? Can leak session keys as well.}

As in~\cite{Norr21}, we verify \mEdhoc{} in the Dolev-Yao model~\cite{DY83}.
%
This is a well-established model for symbolic verification of security protocols.
%
Messages are modelled as terms in an algebra, and the various cryptographic operations are assumed to be perfect, i.e. encrypted messages can only be decrypted with the correct key, no hash collisions exist etc.
%

@@

We now briefly describe \mTamarin{}, and how we model \mEdhoc{} in it here.

%-------------------------------------------------------------------------- sub
\subsection{\mTamarin{}}
\label{sec:tamarin}
As in~\cite{Norr21}, we model and verify \mEdhoc{} in the symbolic model using \mTamarin{}~\cite{DBLP:conf/cav/MeierSCB13}, which is an interactive tool for the formal verification of security protocols.
%
Protocols are modelled in \mTamarin{} as multi-set rewrite rules which encode a transition relation.
%
The elements of these multi-sets, called facts, contribute to the global system state.
%
For syntactic sugar, \mTamarin{} also allows the use of let-bindings and tuples. 
%
For ease of presentation, we will present the model and properties in a slightly different syntax in this paper, but this syntax can be directly mapped to that of \mTamarin.

@@

An annotated multi-set rewrite rule is represented as $l \ifarrow[e] r$, where $l$ and $r$ are multi-sets, and $e$ is a multi-set of actions.
%
A sequence of actions yields a protocol execution trace.
%
Event types are predicates over global states generated during protocol execution.
%
Consider an event type $E$ and a timestamp $t$ as part of a trace.
%
By $E^{t}(p_i)_{i\in\mathbb{N}}$, we represent an event of type $E$ occurring at time $t$ in a trace, parametrized by the sequence of values $(p_i)_{i\in\mathbb{N}}$ (corresponding to the action fact $E(p_i)_{i\in\mathbb{N}}@t$ in \mTamarin).
%
Thus, the time points form a quasi order, and we denote the fact that $t_{1}$ comes before $t_{2}$ in a protocol trace by $t_{1} \lessdot t_{2}$, and that $t_{1}$ and $t_{2}$ stand for the same time point in a trace by $t_{1} \doteq t_{2}$.
%
\mTamarin{} allows events to occur at the same time point, with one restriction: multiple events of the same type cannot occur simultaneously, so if $t_{1} \doteq t_{2}$, then $E^{t_{1}} = E^{t_{2}}$.

Properties are defined as formulas in a fragment of temporal first order logic, and these formulas can be verified over execution traces.

@@

%
The rules modelling a Dolev-Yao adversary and the evolution of their knowledge as the protocol executes are built into \mTamarin.
%
We can extend and/or modify this behaviour by adding other rules as well.
%
To denote that the adversary has access to a message $p$ at time $t$, we use $\mK^{t}(p)$.
%
This stands for the fact that as the adversary interacts with the parties executing the protocol, at time point $t$, the adversary can derive the term $p$ using the Dolev-Yao deduction rules and any additional adversary capabilities specified in \mTamarin.
%
As an example, consider the following implication 
\[
    \forall t, k, k'\mLogicDot \mK^{t}(\langle k, k'\rangle)\ \rightarrow\ 
\mK^{t}(k) \land \mK^{t}(k'),
\]
which represents the fact that if the adversary gets to know the pair of keys $\langle k, k' \rangle$ at a time point $t$, then the adversary also knows each of those keys $k$ and $k'$ at time point $t$ as well.
%
For more details about how \mTamarin{} manages adversary knowledge, see~\cite{DBLP:conf/cav/MeierSCB13}.
%

@@

\subsection{Model and Desired Properties}
\label{sec:desired-properties}
In this section, we describe how we model \mEdhoc{} and the various properties we want it to satisfy.
%

The party $I$ executing the initiator role considers a run of the protocol begun as soon as it sends the first message \mMsgone{} with event type \mIStart, and considers the run ended once it has sent the third message \mMsgthree{} with event type \mIComplete.
%
Similarly, the responder $R$ considers a run started upon receiving \mMsgone{} with event type \mRStart, and finished upon receiving \mMsgthree{} with type \mRComplete.
%

%
In this work, we consider the following properties: perfect forward secrecy, injective agreement for both the initiator and the responder, and implicit agreement. 
%
Agreement is considered on a set of parameters $S$ which also contains the session key material \mSessKey{}.
%
We will first describe in detail both properties, and then describe the contents of the set $S$.
%
We formalize these properties as shown in Figure~\ref{fig:props}, which is a modified version of a figure in~\cite{Norr21}.
\vnote{The figure changes because of TEEs}

\begin{figure*}[ht]
\begin{align*}
    \mPredPfs \triangleq\ & \forall I, R, \mSessKey, t_2, t_3\mLogicDot
    \mK^{t_3}(\mSessKey)\  \land\ 
    (\mIComplete^{t_2}(I, R, \mSessKey)\, \lor\, \mRComplete^{t_2}(I, R, \mSessKey))
    \rightarrow\\
    &(\exists t_1\mLogicDot \mRevLTK^{t_1}(I) \land t_1 \lessdot t_2)
    \ \lor\ (\exists t_1\mLogicDot \mRevLTK^{t_1}(R) \land t_1 \lessdot t_2)
    \ \lor\ (\exists t_1\mLogicDot \mRevEph^{t_1}(R, \mSessKey))
    \ \lor\ (\exists t_1\mLogicDot \mRevEph^{t_1}(I, \mSessKey))
\end{align*}
\begin{align*}
    \mPredInjI \triangleq\ &
    \forall I, R, \mSessKey, S, t_2\mLogicDot \mIComplete^{t_2}(I, R, \mSessKey, S)
    \rightarrow\\
    &(\exists t_1\mLogicDot \mRStart^{t_1}(R, \mSessKey, S) \land t_1 \lessdot t_2)
    \land (\forall I' R' t_1' \mLogicDot \mIComplete^{t_1'}(I' , R', \mSessKey, S)
        \rightarrow t_1' \doteq t_1)
    \ \ \lor\ \ (\exists t_1\mLogicDot \mRevLTK^{t_1}(R) \land t_1 \lessdot t_2)
\end{align*}
\begin{align*}
    \mPredInjR \triangleq\ &
    \forall I, R, \mSessKey, S, t_2\mLogicDot \mRComplete^{t_2}(I, R, \mSessKey, S)
    \rightarrow\\
    &(\exists t_1\mLogicDot \mIStart^{t_1}(I, R, \mSessKey, S) \land t_1 \lessdot t_2)
    \land (\forall I' R' t_1' \mLogicDot \mRComplete^{t_1'}(I' , R', \mSessKey, S)
        \rightarrow t_1' \doteq t_1)
    \ \ \lor\ \ (\exists t_1\mLogicDot \mRevLTK^{t_1}(I) \land t_1 \lessdot t_2).
\end{align*}
\begin{align*}
    \mPredImpI \triangleq\ &
    \forall I, R, \mSessKey, S, t_1\mLogicDot \mIComplete^{t_1}(I, R, \mSessKey, S)
    \rightarrow\\
      &(\forall I', R', S', t_2\mLogicDot \mRComplete^{t_2}(I', R', \mSessKey, S') \rightarrow
             (I=I' \land R=R' \land S=S'))\\
      &\land (\forall I', R', S', t_1'\mLogicDot
        (\mIComplete^{t_1'}(I', R', \mSessKey, S') \rightarrow t_1' \doteq t_1
        )\\
      &\ \ \ \lor(\exists t_0\mLogicDot \mRevLTK^{t_0}(R) \land t_0 \lessdot t_1)
    \lor(\exists t_0\mLogicDot \mRevEph^{t_0}(R, \mSessKey))
    \lor(\exists t_0\mLogicDot \mRevEph^{t_0}(I, \mSessKey)).
\end{align*}
%
\caption{Formalization of security properties and adversary model.}
\label{fig:props}
\end{figure*}

%----------------------------------------------------------------------- PFS
\subsubsection{Perfect Forward Secrecy (PFS): }
\label{sec:secrecy}
Informally, by perfect forward secrecy, we mean that the adversary cannot gain access to the session key material \mSessKey{}, even if a long-term key is leaked after the establishment of \mSessKey{}.
%
PFS for session key material \mSessKey{} is formalized as \mPredPfs{} in Figure~\ref{fig:props}.

In the \mIComplete{} event, $I$ represents the identity of the initiator, and $R$ represents the identity of the party who the initiator believes is playing the responder role, while \mSessKey{} stands for the established session key material.
%
\mRComplete{} has analogous parameters; in particular, the responder $R$ believes $I$ is the party playing the initiator role.
%
Essentially, the property states that an adversary obtains \mSessKey{} only if a party's long-term key is compromised before their run ends, or if an ephemeral key is compromised at any time after a party's run began.
\vnote{This has to change once we add TEEs}
%
The property in Figure~\ref{fig:props} is unlike the actual \mTamarin{} lemma in one minor way: \mTamarin's syntax does not allow disjunctions to appear on the left-hand side of an implication inside a universally-quantified formula.
%
Therefore, in the \mTamarin{} code, instead of using the disjunction $\mIComplete^{t_2}(I, R, \mSessKey)\, \lor\,  \mRComplete^{t_2}(I, R, \mSessKey)$ to model the fact that either party can have completed their execution, we use a single action parametrized by the terms $I$, $R$, and \mSessKey.

%-------------------------------------------------------------------- InjAgree
\subsubsection{Authentication Properties:}
\label{sec:authenticationDef}
Following~\cite{Norr21}, we prove two different kinds of authentication properties, namely \emph{injective agreement} in the style of~\cite{DBLP:conf/csfw/Lowe97a}, and implicit agreement.
%
Injective agreement can be guaranteed to either party running the protocol.
%
For the initiator $I$, it guarantees to $I$ that whenever $I$ believes that they have completed a run with $R$ as responder, then the party $R$ has indeed executed the protocol in the role of a responder, and that this run of $I$ uniquely corresponds to one of $R$ where the set of parameters is $S$ and includes, in particular, the session key material \mSessKey{}.
%
It can be defined for $R$ in a similar manner.

%
We formalize injective agreement for the initiator role as \mPredInjI{} and for the responder role as \mPredInjR{} in Figure~\ref{fig:props}.
%
For the initiator $I$, this property represents the fact that either injective agreement (as described above) holds, or the long-term key of the party $R$ assumed to be playing the responder role has been compromised before $I$'s role finished.
%
If the adversary has managed to compromise $R$'s long-term key, they can generate any message of their liking and signed it using said key, or computed a $\mathit{MAC}_{R}$ using the long-term keys \mPubE{I}, \mPriv{R}, and an ephemeral key pair $\langle\mPrivE{R},\ \mPubE{R}\rangle$ of their choice.
%
The analogous definition holds for the responder $R$.
\vnote{Check this, and modify to deal with TEEs}

%------------------------------------------------------------- Implicit auth
As part of our analysis, we found that all the \mEdhoc{} methods satisfy PFS.
%
However, this is not the case for the injective agreement property as stated above.
%
Thus, we show a different property, a form of implicit agreement on the same set of parameters, which is guaranteed for all methods.
%
This modification is inspired by the definitions of implicit authentication in the computational model~\cite{DBLP:conf/csfw/GuilhemFW20}.
%
While that paper focuses on authenticating just a key and related identities, our definition encompasses a general set of parameters, as in the notion of injective agreement proposed by Lowe~\cite{DBLP:conf/csfw/Lowe97a}.

The ``implicit'' in the name of the property stands for the fact that a party $A$ assumes that any party $B$ who has access to the session key material \mSessKey{} must, in fact, be the intended party, and that if $B$ is honest, $B$ will agree on a set $S$ of parameters which includes \mSessKey.
%
Implicit agreement for both roles guarantees to $A$ that $A$ is or has been involved in exactly one protocol execution with $B$, and that $B$ agrees or will agree with $A$ on $S$.
%
This property diverges from injective agreement in that upon sending the last message, $A$ concludes that if this message reaches $B$, then $A$ and $B$ agree on each other's identities and roles, and the set $S$.

Note that for implicit agreement to hold for $I$, the ephemeral keys must not be revealed, since the property relies on the fact that the intended responder is the only one who knows the session key material.
%
If the adversary has access to the ephemeral keys, they can use it along with the public keys of $I$ and $R$ to compute the session key material.
%
However, either party's long-term key can be revealed after the other party has finished their execution, since this still leaves the adversary unable to compute \mGxy{}.

Since \mTamarin{} runs out of memory to verify this property as is, we split it into two lemmas -- \mPredImpI{} for $I$ for one \mPredImpR{} for $R$.
%
Figure~\ref{fig:props} contains the definition for \mPredImpI{}.
%
\mPredImpR{} is formalized similarly, so we omit it.

%------------------------------------------------------- Agreed parameters
\subsubsection{Parameters in set $S$:}
\label{sec:agreedParams}
We now describe the set $S$ of parameters upon which the two parties obtain guarantees via the above properties.
%
The initiator $I$ gets injective and implicit agreement guarantees on the
following partial set $S_P$ of parameters~\cite{Norr21}:
\begin{itemize}
    \item the roles played by itself and its peer,
    \item responder identity,
    \item session key material (which varies depending on the \mEdhoc{} method),
    \item context identifiers \mCi{} and \mCr{}, and
    \item cipher suites \mSuites{}.
\end{itemize}
%

Due to the initiator being guaranteed identity protection under \mEdhoc{}, $I$ cannot get explicit agreement with $R$ on the initiator's identity.
%
Similarly, when using the \mStat{} authentication method, $I$ does not get any such guarantees about $P_{I}$.
%
However, $I$ does get implicit agreement with $R$ about $I$'s identity and the full set $S_{F}$ of agreed parameters.
%
In contrast, since $R$'s run finishes after that of $I$, $R$ can get explicit injective agreement assurances on the full set $S_{F}$ of agreed parameters.
%
The full set of agreed parameters $S_F$ is $S_P \cup \{I, P_I\}$ when $P_I$ is part of the session key material, and $S_P \cup \{I\}$ otherwise.
%

In addition to these properties, a couple of properties can be inferred without being explicitly modelled and verified. 
%
One might be interested in checking is called Key-Compromise Impersonation (KCI)~\cite{DBLP:conf/ima/Blake-WilsonJM97}.
%
Key-Compromise Impersonation (KCI)~\cite{DBLP:conf/ima/Blake-WilsonJM97} often occurs in protocols where a party cannot be assured that the peer they are involved in a run with actually knows the session key material.
%
A KCI attack occurs when an adversary who has access to $A$'s long-term private key to make $A$ believe that they completed an execution with a peer $B$, while $B$ did not participate in said execution at all.
%
Our above notions of agreement ensure that both parties agree on each other's identity, role, and session key material.
%
Therefore, all \mEdhoc{} methods that satisfy these agreement properties also avoid KCI attacks.

Another kind of attack is an Unknown Key Share (UKS)~\cite{DBLP:conf/ima/Blake-WilsonJM97}.
%
As part of a UKS attack, a party $A$ can be made to believe that it finished an execution with a party $B$, but where the session key material is actually shared between $A$ and $C$ instead.
%
Again, due to the agreement on identities and session key material, any method that satisfies the above agreement properties also resists UKS attacks.
%
Overall, the agreement properties capture entity authentication, and satisfy any properties based on that notion.

\vnote{Is there anything else we can add thanks to the TEEs?}

%---------------------------------------------------------------------------sub
\subsection{Encoding \mEdhoc{} in \mTamarin}
\label{sec:modeling}
%
In this section, we describe how we model the \mSigSig, \mSigStat, \mStatSig, and \mStatStat{} methods of \mEdhoc{} in \mTamarin.
%
As in~\cite{Norr21}, we construct the \mTamarin{} model by utilizing the fact that all methods of \mEdhoc{} share a common underlying structure (as shown in Figure~\ref{fig:edhocFramework}).
%
We do so by writing a single specification file in the M4 macro language, which generates all the methods.
%
As in~\cite{Norr21}, we only present the \mStatSig{} method which illustrates the use of two different authentication methods.
%
The full \mTamarin{} code for all models can be found at~\cite{edhocTamarinRepo}.
%

As mentioned in Section~\ref{sec:tamarin}, we extend the default equational theory of \mTamarin{} to handle various operations used in \mEdhoc.
%
Of the built-in theories, we use the ones for exclusive-or (\mXor), Diffie-Hellman operations, signatures (\mT{sign} and \mT{verify} operations), and hashing~\cite{DBLP:conf/csfw/DreierHRS18,DBLP:conf/csfw/SchmidtMCB12}.
%
We augment the hashing function symbol to add an extra input, which we use to model different hash functions.
%

In addition to these default operations, \mEdhoc{} is built over \mHkdfExpand, \mHkdfExtract, and the \mAead{} functions.
%
We represent \mHkdfExpand by \mT{expa} and \mHkdfExtract by \mT{extr}.
%
\mAead{} operations need us to add extra equations to the underlying theory.
%
A term encrypted using \mAead{} is represented by \mT{aeadEncrypt(m, k, ad, ai)}, where \mT{m} is the underlying message, \mT{k} is the encrypting key, \mT{ad} is the additional data, and \mT{ai} is the identifier for the encryption algorithm.
%
Decryption of such a term is defined via two equations that we add to \mTamarin's theory.
%
The following equation requires the decrypting party to know the additional data \mT{ad} to decrypt this encrypted term with verification of its integrity. 
\begin{small}
\begin{verbatim}
aeadDecrypt(aeadEncrypt(m, k, ad, ai), k, ad, ai) = m.
\end{verbatim}
\end{small}
%
Only the above equation is used by honest parties, but the adversary should also be able to decrypt without having to go through the additional step of identity verification.
%
To this end, we also add the following equation, where the adversary does not need access to the additional data \mT{ad} in order to decrypt.
\begin{small}\begin{verbatim}
decrypt(aeadEncrypt(m, k, ad, ai), k, ai) = m.
\end{verbatim}\end{small}
%

Having described how we adapt the equational theory to model \mEdhoc, we now move on to the modelling of the adversary model and the environment in which the protocol is executed.
%
We extend the built-in Dolev-Yao adversary rules which are part of \mTamarin.
%
We use the following rules to capture the link between a party's identity and their long-term key pairs, in the \mSig{}- and in the \mStat{}-based methods respectively.
\begin{center}
\begin{minipage}{0.48\textwidth}
\begin{small}
\begin{verbatim}
1 rule registerLTK_SIG:
2    [Fr(~ltk)] --[UniqLTK($A, ~ltk)]->
3        [!LTK_SIG($A, ~ltk),
4         !PK_SIG($A, pk(~ltk)),
5         Out(<$A, pk(~ltk)>)]
\end{verbatim}
\end{small}
\end{minipage}
\hfill\vline\hfill
\begin{minipage}{0.48\textwidth}
\begin{small}
\begin{verbatim}
1 rule registerLTK_STAT:
2    [Fr(~ltk)] --[UniqLTK($A, ~ltk)]->
3        [!LTK_STAT($A, ~ltk),
4         !PK_STAT($A, 'g'^~ltk),
5         Out(<$A, 'g'^~ltk>)]
\end{verbatim}
\end{small}
\end{minipage}
\end{center}

Using the fact \verb|Fr(~ltk)|, \mTamarin{} creates a new term \mT{ltk} and uses it to represent a secret long-term key.
%
Via the \verb|Out(<$A, pk(~ltk)>)| fact, \mTamarin{} puts out onto the communication channel the identity of the party to whom this long-term key belongs, along with their public key. 
%
Since the adversary has access to the communication channel, they can pick up all of this information.
%
The event \mT{UniqLTK} parametrized by a party's identity and their long-term key models the unique correspondence between those two values.
%
As a result, this rules out a party owning multiple long-term keys -- in particular, it keeps the adversary from registering long-term keys in some honest party's name.
%
This aligns well with an external mechanism such as a certificate authority ensuring that long-term keys are uniquely assigned to the corresponding identities, which is ensured by the \mSpec.
